---
title: "Exercise 3 pdf"
output: pdf_document
---

```{r eval = FALSE}
install.packages('arrow')
install.packages('wru')
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 5)

library(tidyverse)
library(lubridate)
library(arrow)
```

## Load data

Load the following data:
  + applications from `app_data_sample.parquet`
  + edges from `edges_sample.csv`

```{r load-data}
# change to your own path!
applications <- read_parquet("app_data_sample.parquet")
edges <- read_csv("edges_sample.csv")
applications
edges
```

## Get gender for examiners

We'll get gender based on the first name of the examiner, which is recorded in the field `examiner_name_first`. We'll use library `gender` for that, relying on a modified version of their own [example](https://cran.r-project.org/web/packages/gender/vignettes/predicting-gender.html).

Note that there are over 2 million records in the applications table -- that's because there are many records for each examiner, as many as the number of applications that examiner worked on during this time frame. Our first step therefore is to get all *unique* names in a separate list `examiner_names`. We will then guess gender for each one and will join this table back to the original dataset. So, let's get names without repetition:

```{r gender-1}
library(gender)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it
# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)
examiner_names
```

Now let's use function `gender()` as shown in the example for the package to attach a gender and probability to each name and put the results into the table `examiner_names_gender`

```{r gender-2}
# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
examiner_names_gender
```

Finally, let's join that table back to our original applications data and discard the temporary tables we have just created to reduce clutter in our environment.

```{r gender-3}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)
# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")
# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```


## Guess the examiner's race

We'll now use package `wru` to estimate likely race of an examiner. Just like with gender, we'll get a list of unique names first, only now we are using surnames.

```{r race-1}
library(wru)
examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()
examiner_surnames
```
We'll follow the instructions for the package outlined here [https://github.com/kosukeimai/wru](https://github.com/kosukeimai/wru).

```{r race-2}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race
```

As you can see, we get probabilities across five broad US Census categories: white, black, Hispanic, Asian and other. (Some of you may correctly point out that Hispanic is not a race category in the US Census, but these are the limitations of this package.)

Our final step here is to pick the race category that has the highest probability for each last name and then join the table back to the main applications table. See this example for comparing values across columns: [https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/). And this one for `case_when()` function: [https://dplyr.tidyverse.org/reference/case_when.html](https://dplyr.tidyverse.org/reference/case_when.html).

```{r race-3}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
examiner_race
```

Let's join the data back to the applications table.

```{r race-4}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)
applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))
rm(examiner_race)
rm(examiner_surnames)
gc()


# applications = applications %>% rename(race = race.x, gender = gender.x)

# applications = applications %>% select( -c(race.y, gender.y))

names(applications)
```


## Examiner's tenure 

To figure out the timespan for which we observe each examiner in the applications data, let's find the first and the last observed date for each examiner. We'll first get examiner IDs and application dates in a separate table, for ease of manipulation. We'll keep examiner ID (the field `examiner_id`), and earliest and latest dates for each application (`filing_date` and `appl_status_date` respectively). We'll use functions in package `lubridate` to work with date and time values.

```{r tenure-1}
library(lubridate) # to work with dates
examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 
examiner_dates
```

The dates look inconsistent in terms of formatting. Let's make them consistent. We'll create new variables `start_date` and `end_date`.

```{r tenure-2}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

Let's now identify the earliest and the latest date for each examiner and calculate the difference in days, which is their tenure in the organization.

```{r tenure-3}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
    filter(year(latest_date)<2018)
examiner_dates
```

Joining back to the applications data.

```{r tenure-4}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")
rm(examiner_dates)
gc()
```
```{r}

save(applications,file="applications.Rda")
save(edges,file="edges")
```


## Network Analysis 
After running the code to generate the 3 new features: gender, race and tenure, we proceed to analyze the connections in the network by sampling 2 workgroups.

```{r}

load(file='applications.Rda')

load(file='edges')



names(applications)
names(edges)



applications %>% count(examiner_art_unit, sort = TRUE)
```

```{r plots1}

# Choosing workgroups 176 and 179
wg1 = applications %>% filter(substr(examiner_art_unit, 1, 3) == '176' ) %>% 
  arrange(application_number)

wg2 = applications %>% filter(substr(examiner_art_unit, 1, 3) == '179' ) %>% 
  arrange(application_number)

#summary(wg1)


# distributions for wg 176
p1 = wg1 %>% group_by(race) %>% summarise(n_examiners = n_distinct(examiner_id)) %>% 
  ggplot(aes(x = race, y = n_examiners)) + geom_bar(stat ='identity')

p2 = wg1 %>% group_by(gender) %>% summarise(n_examiners = n_distinct(examiner_id)) %>% 
  ggplot(aes(x = gender, y = n_examiners)) + geom_bar(stat ='identity')

p3 = wg1 %>% ggplot(aes(x = tenure_days)) + geom_histogram()

# library(gridExtra)
# par(mfrow=c(1,3))
# grid.arrange(p1, p2, p3, ncol=3)
print(p1)

print(p2)

print(p3)



```

```{r}
# distributions for wg 179
p1 = wg2 %>% group_by(race) %>% summarise(n_examiners = n_distinct(examiner_id)) %>% 
  ggplot(aes(x = race, y = n_examiners)) + geom_bar(stat ='identity')

p2 = wg2 %>% group_by(gender) %>% summarise(n_examiners = n_distinct(examiner_id)) %>% 
  ggplot(aes(x = gender, y = n_examiners)) + geom_bar(stat ='identity')

p3 = wg2 %>% ggplot(aes(x = tenure_days)) + geom_histogram()

# par(mfrow=c(1,3))
# grid.arrange(p1, p2, p3, ncol=3)
p1

p2

p3
```


## Creating advice networks

```{r}

#finding edges of first workgroup. We keep the edges of applications that belong to wg1
edges_wg1 = edges %>% inner_join(wg1[c('application_number')] , by = 'application_number') 
edges_wg1  = drop_na(edges_wg1)

names(edges_wg1)

dim(edges_wg1)

#creating the network object
library(tidygraph)
library(tidyverse)
library(ggraph)

edges_wg1 = edges_wg1 %>% rename(to = alter_examiner_id ,
                     from = ego_examiner_id )

graph = as_tbl_graph(x = edges_wg1[c('to','from')])

graph = graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree())

ggraph(graph, layout = 'graphopt') + 
  geom_edge_link()+
  geom_node_point(aes(size = centrality_degree, colour = centrality_degree))+
 theme_graph()

nodes_df = graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree()) %>%  data.frame()


summary(nodes_df$centrality_degree)

```


```{r}

nodes_df


wg1$examiner_id = as.character(wg1$examiner_id)

wg1_race_centrality = nodes_df %>% left_join(wg1, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, race) %>% distinct() %>% 
  group_by(race) %>% summarise(mean_degree = mean(centrality_degree), 
                               n_examiners = n()) 

wg1_race_centrality

wg1_race_centrality %>% filter(race %in% c('Asian', 'white')) %>% ggplot(aes(x = race, y = n_examiners, fill = mean_degree)) +
geom_bar(stat = 'identity')
```


Even though there are more white members in the network, they have, on average, less connections (or advice exchanges) than the second most common race.

```{r}


wg1_gender_centrality = nodes_df %>% left_join(wg1, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, gender) %>% distinct() %>% 
  group_by(gender) %>% summarise(mean_degree = mean(centrality_degree), 
                               n_examiners = n()) 


wg1_gender_centrality

wg1_gender_centrality %>% filter(gender %in% c('female', 'male')) %>%  ggplot(aes(x = gender, y = n_examiners, fill = mean_degree)) +
geom_bar(stat = 'identity')


```
Similar to the gender analysis: underrepresented groups tend to "overcompensate" by having more exchanges than the more frequent groups. In this case we see that females tend to have 4 times the connections as males when it cames to asking or receiving advice.

```{r}

wg1_tenure_centrality = nodes_df %>% left_join(wg1, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, tenure_days) %>% distinct()

wg1_tenure_centrality %>% ggplot(aes(x = centrality_degree, y = tenure_days)) + geom_point() +
  geom_smooth(method='lm') + xlab('centrality degree') + ylab('tenure days') + ggtitle('Scatterplot: centrality degree and tenure days')

```
This scatter plot shows that examiners that have a shorter tenure in the organization tend to ask/give advice more times compared to the more seasoned examiners. One possible explanation is that novice examiners tend to ask for more feedback of their work. Even though some senior examiners will have a high degree centrality, they are too few compared to the big number of junior examiners, so the regression is tilted towards the latter and ends up showing a negative relationship. 


```{r}

#finding edges of first workgroup. We keep the edges of applications that belong to wg1
edges_wg2 = edges %>% inner_join(wg2[c('application_number')] , by = 'application_number') 
edges_wg2  = drop_na(edges_wg2)

#creating the network object
library(tidygraph)
library(tidyverse)
library(ggraph)

edges_wg2 = edges_wg2 %>% rename(to = alter_examiner_id ,
                     from = ego_examiner_id )

graph = as_tbl_graph(x = edges_wg2[c('to','from')])

graph = graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree())

ggraph(graph, layout = 'graphopt') + 
  geom_edge_link()+
  geom_node_point(aes(size = centrality_degree, colour = centrality_degree))+
 theme_graph()

nodes_df = graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree()) %>%  data.frame()


summary(nodes_df$centrality_degree)


```


```{r}

wg2$examiner_id = as.character(wg2$examiner_id)

wg2_race_centrality = nodes_df %>% left_join(wg2, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, race) %>% distinct() %>% 
  group_by(race) %>% summarise(mean_degree = mean(centrality_degree), 
                               n_examiners = n()) 

wg1_race_centrality

wg1_race_centrality %>% ggplot(aes(x = race, y = n_examiners, fill = mean_degree)) +
geom_bar(stat = 'identity')
#filter(race %in% c('Asian', 'white')) %>% 
```

```{r}

wg2_gender_centrality = nodes_df %>% left_join(wg2, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, gender) %>% distinct() %>% 
  group_by(gender) %>% summarise(mean_degree = mean(centrality_degree), 
                               n_examiners = n()) 


wg2_gender_centrality

wg2_gender_centrality %>% filter(gender %in% c('female', 'male')) %>%  ggplot(aes(x = gender, y = n_examiners, fill = mean_degree)) +
geom_bar(stat = 'identity')
```
```{r}


wg2_tenure_centrality = nodes_df %>% left_join(wg2, by=c('name' = 'examiner_id')) %>% 
  select(name, centrality_degree, tenure_days) %>% distinct()

wg2_tenure_centrality %>% ggplot(aes(x = centrality_degree, y = tenure_days)) + geom_point() +
  geom_smooth(method='lm') + xlab('centrality degree') + ylab('tenure days') + ggtitle('Scatterplot: centrality degree and tenure days')

```
A much more "connected" network with higher centrality degrees in average (4 vs 2 when taking average of degree) 


