---
title: "Exercise 2"
output: pdf_document
---

```{r}
library(dplyr)
library(tidygraph)
library(tidyverse)
library(ggraph)
```

```{r}
df = read.csv('Fakebook_edges.csv')
df
```
```{r}

graph = as_tbl_graph(df)

#https://www.data-imaginist.com/2017/introducing-tidygraph/
# printing the network
ggraph(graph) + 
  geom_edge_link() + 
  geom_node_point()+
  geom_node_text(aes(label = name)) +
  theme_graph()

```
```{r}

# centrality degree: counting edges of each note
graph =graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree())

graph
```


```{r}
# centrality degree: counting edges of each note
graph = graph %>%
  activate(nodes) %>%
  mutate(centrality_betweenness = centrality_betweenness())
```
 
```{r}
graph %>% 
  filter(name %in% c(7,8,9,10))
```
Based on the computations of degree and betweenness centralities, we can see that spots B (8) and A (7) are the most appealing to sit in. First, between spots B,C,D, they all share the same centrality degree of 5 (5 edges each), so the thing that sets them apart is the centrality betweenness --how valuable is their position to bring together separate 'cliques' in the network--. From there we can see that B is the preferable spot. Going further, we see that spot A (7) has a notably lower degree centrality, but however outshines the rest in terms of betweenness score. 

Therefore, it will depend more on the intern and his/her intentions or preferences for relationship building to decide which of these 2 to select. In my opinion, considering an internship of 3 months, I think it is better to maximize the amount of trusting relationships built, so I lean towards choosing seat B.
